---
title: "A reflection on Shiny"
description: |
  Recently i decided to put my Shiny skill to the test with a mini project that took one week to complete. Now it's time to reflect on what i have done well and what can be improved. Hope you will find something interesting out of what i have to say!
author:
  - name: Hanh Ngo
    affiliation: Monash University
    program: Master of Business Analytics
date: 10-16-2020
output:
  distill::distill_article:
    self_contained: false
---

Hello again!

It's this time of month where i will emerge and pop out something new on this blogpost. So last week I was attempted to create something a little more complex than what i usually do. Its original purpose was for a school project, but i saw this as quite a challenge to what i was accustomed to, which used to be markdowns and some cute and ready-for-running HTML pages knitted from those markdowns. It was a shiny app with a whole lots of different syntax, having different structure and holding a handful of new definition. The topic for this Shiny app was the same old Covid-19 that people cannot seem to stop talking about these days. I might want to see if i can create something new out of the same old stuff.  

# About this Shiny app i keep talking about
The Shiny app i build this time named "The Vietnam COVID-19". As the named implied, this app was built focusing on reporting the COVID situation in Vietnam from the time when the pandemic first outbroke in the country. I chose Vietnam simply because this is my country so i probably had a lot to tell (I tried).  

The app was sectioned into four subsection, which are:  

- The _Bulletin Board_ page which provided some recaps on the current situation of the country. Information like the total confirmed cases, total recovered and total death can be found here. A bar chart showing change of new cases by day was also featured here, along with the list of confirmed cases which can be unfolded by clicking on the day of your choice in the bar chart.  

```{r include-dashboard}
knitr::include_graphics(file.path(getwd(),"dashboard.png"))
```


- In the _Infection Route_ page, you will be given a handful of information. We started off by having a network on the left hand side where you can review the infection route of patients for each month. You are in the control of which legend to show on the chart by toggling on and off the legend of your choice. On the right hand side, we have the information of each patient featured on the chart for your selected month. In summary, you can learn of the total cases for each month, the confirmed connection between cases, breakdown of numbers by status (recovered, active, deceased) and the patient information featured for your reference.  

```{r include-infectionroute}
knitr::include_graphics(file.path(getwd(),"infection-route.png"))
```

- The _World Situation_ will give you a trip around the world to see how other countries are doing. In this page, you will explore some basic COVID-19 stats for all of the countries in the world, along with a breakdown to details by each country of your choice by selecting the input in the right hand side panel and scrolling down to the table underneath the grand map. 

```{r}
knitr::include_graphics(file.path(getwd(),"world-situation.png"))
```

- Finally, the _About_ will be the place where i gave my acnowledgment and reference to other people's works and resources that inspired me during my creation of this app. A nice image of commemory stamps issued in Vietnam to encourage the health workers force was also featured so it could sooth your hurting eyes from the jungle of words in this page.  

```{r}
knitr::include_graphics(file.path(getwd(),"about.png"))
```

# What i have done well (or at least i believe i have done well)
– What did you do well?


# What i could have done better

– What are the points you can improve on?


– What are some things that you’d like to add that’s plausible?

Translation - More information
Automatic update


– What are some things that you’d like to add that cannot be done yet due to limitation in data or technology?

- color for the value box
- legend colour



