---
title: "The current state of a kingdom"
description: |
  Let's dwelve into our first blog by doing a small visualisation exercise of the covid situation in Israel! I intend to make this first blog to be less formal but more friendly, so we can get to know each other better. Let's visualize together!
author:
  - name: Hanh Ngo
    ID: 31196101
    affiliation: Monash University
    program: Master of Business Analytics
date: 09-12-2020
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
library(tidyverse)
library(shiny)
library(leaflet)
library(htmlwidgets)
library(htmltools)
library(geojsonio)
library(geojsonlint)
library(DT)
library(RColorBrewer)
library(gganimate)
library(ggplot2)
library(kableExtra)
library(lubridate)
library(plotly)
library(gifski)
library(png)
library(magick)
```

# Introduction
Hello! and welcome to the second blog post of "New things per day" series. In this blog post, we will explore the 

Choose this country purposely to show the counter intuitive

# Data description


# Everything you need to know about Iraq
Irad used to be regarded as the "cradle of civilization" of the Middle Ages. The country is a home to many heritage and historic sites, originating back from the golden time of the Ottoman regime. Some of the cultural heritage we may have heard of would include the Hanging garden in Babylon, the remnants of palace and buildings.....

However, looking at the country now, all we see is a turbulence of time of wars, fightings and violences. Hundreds of Iraq people fled the country to run away from the terrorists, people died and lost their homes amid the bombing .... (find reference for this).

In the next section, i will try to recap some of the highlights of the Iraq's contemporary current situation based on the choronological order. I handpicked only events that i believed directly led to the current state of Iraq. Before I start, i want you to answer this question yourself: "How do you make of Iraq? Is it a poor country torned by wars with high children mortality rate, poor facility, high iliteracy rate and ect..". Please keep your answer in mind and proceed. 

## The history starts
Our story started 

## Kurdish Internal conflict
Kurds have never achieved nation-state status, making Kurdistan a non-governmental region and one of the largest stateless nations in the world.  
Portions of the region are recognized by two countries: Iran, where the province of Kordestan lies; and northern Iraq, site of the autonomous region known as Kurdistan Regional Government (KRG) or Iraqi Kurdistan.  
Kurds were mostly nomadic until the end of World War I and the breakup of the Ottoman Empire.  
Kurds make up about 10% of the population in Syria, 19% of the population of Turkey, 15-20% of the population of Iraq and are the second largest ethnicity in Iran.  

```{r kurdis-timeline, out.width= '100%'}

```


## Terrorists

As of today, Iraq is still torned with many terrorists activities. The Central Intelligence agency (CIA) of America has pointed out that the country currently hosts 5 home- based terrorist groups home and 2 foreign - based terrorist groups, one of which you might have heard a lot in the media - the Islamic State of Iraq and ash-Sham (ISIS). With the   

The Armed Conflict Location & Event Data Project (ACLED) of the United States has tried to put together all of the political violence events and some relevant non - relevant events happened everyday around the world. I obtained a list for Iraq from their database, which can be downloaded [here](https://acleddata.com/data-export-tool/), for the period of 2016 - the most recently of 2020. Using this dataset, i tried to visualise the spread of the terrorist attacks in the territory of Iraq, showing how the country have suffered from the acts for terrorism.

First we have a look at the dataset provided by the ACLED: 

```{r}
terror <- read.csv(here::here("Blog_entry_1/data", "iraq_terrorists.csv")) %>%
  mutate(start = as.Date(as.character(start), format = "%d-%b-%y"),
         end = as.Date(as.character(end), format = "%d-%b-%y")) %>%
  # trick to show the markers on the map. If end = start, no markers will be shown.
  mutate(end = start + 2)

# Choose violent events only for the sake of this report
terror <- terror %>%
  filter(event_type %in% c("Battles", "Explosions/Remote violence", "Violence against civilians")) %>%
  # create group for IS and non-IS events
  mutate(group = ifelse(actor1== "Islamic State (Iraq)" |actor2 == "Islamic State (Iraq)", "IS", "non-IS"),
         color = ifelse(group == "IS", "red", "blue"))

glimpse(terror)
```

ACLED divided the event types into three major groups: Violent events, Demonstrations (which includes protests and riots) and Non - violent actions. A total of 26,235 events were recorded from 2016 to 2020. For the purpose of this section, i filtered only the Violent events to get 21,511 events fitting into the category, which made up a whooping c.82% of total events recorded!

I only used some of the variables featured in the dataset for this section. Also i created some more to facilitate my construction of the map, they are:  

- start: the day of the event.  
- end: i created this variable by taking start day plus 2. This variable was made only for the plotting purpose.  
- event_type: the Violent events are divided into three subgroups: Battles, Explosions/ Remote violence and Violence against vilians.    
- sub_event_type: smaller subgroups under each subgroup mentioned in the event_type to describe the main activities involved in the event.  
- actor1 and actor2: parties involved in the event.  
- location: loction that the events took place.  
- latitude and longitude: the GPS coordinates of the events' location. 
- notes: the detailed description of the events.  
- fatalities: the recorded number of the fatalitites. Eventhough this is an information variable, the data was not collected enough for analysis purpose.  
- group: i created this variable to put all of the events into two groups: IS- related and non-IS-related.
- color: color specifies for each group. This is for plotting purpose only.

Next up, we look at the animated map of all violence events that took place in Iraq from 2016 to 2020:  

```{r violences-plot, fig.cap="The distribution of terrorists events in Iraq from 2016 to 2020", out.width= "100%"}
# get the coordinates for Iraq
download.file(url = 'http://biogeo.ucdavis.edu/data/diva/adm/IRQ_adm.zip', 
              destfile = 'iraq.zip')
unzip(zipfile = 'iraq.zip')

#Read the shp file
library(rgdal)
iraq <- readOGR('IRQ_adm0.shp')

# use geojsonio to convert our data.frame
#  to GeoJSON which timeline expects
violence_geo <- geojson_json(terror,lat="latitude",lon="longitude")

# create a leaflet map on which we will build
library(leaftime)
leaflet(violence_geo) %>%
  addTiles() %>%
  addPolygons(data = iraq, opacity = 0.5, fillOpacity = 0.2, color = "pink") %>%
  setView(42, 32, 5) %>%
  addLegend(colors = unique(terror$color),
            position = "topright",
            labels = unique(terror$group), 
            title = "Responsible by",
            opacity = 1) %>%
  addTimeline(
    sliderOpts = sliderOptions(
    formatOutput = htmlwidgets::JS(
               "function(date) {return new Date(date).toLocaleDateString()}
               "),
             position = "bottomleft",
             step = 1000,
             duration = 100000,
             showTicks = TRUE
             ),
    timelineOpts = timelineOptions(
      pointToLayer = htmlwidgets::JS(
    "
    function(data, latlng) {
      return L.circleMarker(latlng, {
        radius: 50,
        color: data.properties.color,
        fillColor: data.properties.color,
        fillOpacity: 0.25
      }).bindTooltip(
      '<strong/>Date: </strong>'+ data.properties.start + 
      '<br/><strong/>Event: </strong>' +data.properties.event_type + 
      '<br/><strong/>Sub-event: </strong>' +data.properties.sub_event_type +
      '<br/><strong/>Parties: </strong>' +data.properties.actor1 + ' vs '+ data.properties.actor2+
      '<br/><strong/>Location: </strong>' +data.properties.location,
      {permanent: false},
      ).openTooltip();
    }
    "
  )
))
```

You can hit the play button in the slider to animate the event map by times. You can also use the forward and backward buttons to control the apprearance of events. By hovering your cursor over the points, you can read the description for each events shown.   

Most violent events in Iraq took place in the North East of the country, cluttering around the Kurdistan and the central regions. We can see that a majority of violent events in Iraq came from the ISIS with the red points dominant the map. The density of the events was particularly dense throughout 2016 but situation started to lift up in 2017 as the Iraq Government launched a campaign to take back Mosul - the second biggest city of Iraq the last remaining ISIS stronghold in Iraq in October 2016. By July 2017, Mosul was liberated. By December 2017, the Iraqi militaries announced that the Iraq's territory was "fully liberated" form the ISIS. From 2018 onwards, we can see that the activities of ISIS were greatly diminished in number in Iraq and the operational area focused mainly on the rural and desert areas. 

This animated map time was constructed using the magnificent works of _timelyportfolio_ on his **leaftime** package. You can visit his github page for the leaftime package [here](https://github.com/timelyportfolio/leaftime) and explore the usage of this wonderful package.  


## Foreign relations:

### The relation with Iran

1980-1988 - Iran-Iraq war results in stalemate.
1981 June - Israeli air raid destroys Iraqi nuclear reactor at Osirak near Baghdad.
1988 March - Iraq attacks Kurdish town of Halabjah with poison gas, killing thousands.

### The relation with US

1990 - Iraq invades and annexes Kuwait, prompting what becomes known as the first Gulf War. A massive US-led military campaign forces Iraq to withdraw in February 1991.  
1991 April - Iraq subjected to weapons inspection programme.  
1991 Mid-March/early April - Southern Shia and northern Kurdish populations - encouraged by Iraq's defeat in Kuwait - rebel, prompting a brutal crackdown.  
1991 April - UN-approved haven established in northern Iraq to protect the Kurds. Iraq ordered to end all military activity in the area.  
1992 August - A no-fly zone, which Iraqi planes are not allowed to enter, is set up in southern Iraq.  
1995 April - UN allows partial resumption of Iraq's oil exports to buy food and medicine in an oil-for-food programme.  
1996 September - US extends northern limit of southern no-fly zone to just south of Baghdad.  
1998 October - Iraq ends cooperation with UN Special Commission to Oversee the Destruction of Iraq's Weapons of Mass Destruction.  

## The Covid-19

```{r data-load}
data <- read.csv("https://opendata.ecdc.europa.eu/covid19/casedistribution/csv", na.strings = "", fileEncoding = "UTF-8-BOM")
raw_iraq <- data %>%
  filter(countriesAndTerritories == "Iraq") 
```

```{r raw-data}
raw_dat <- raw_iraq %>%
  select(dateRep, day, month, year, cases, deaths, popData2019) %>%
  arrange(year, month, day) %>%
  mutate(cumulative_case = cumsum(cases),
         cumulative_death = cumsum(deaths)) %>%
  mutate(dateRep = dmy(dateRep))
```

```{r fig.cap = "COVID-19 in Iraq"}
p <- raw_dat %>%
  ggplot() +
  geom_line(aes(dateRep, cumulative_case), color = "blue") +
  geom_line(aes(dateRep, cumulative_death), color = "red") +
  theme_classic() +
  ylab("") +
  xlab("Date") +
  theme(axis.text.x=element_text(angle=90, hjust=1)) +
  scale_x_date(breaks = seq.Date(from = as.Date("2019-12-31"), 
                                     to = as.Date("2020-09-12"), by = 7)) +
  scale_y_continuous(labels = scales::comma,
                     breaks = seq(from = 0, to = 300000, by = 25000)) +
  ggtitle("COVID-19 in Iraq") +
  annotate("text", x = as.Date("2020-08-04"), y = 275000, label = "Cumulative cases") +
  annotate("text", x = as.Date("2020-08-11"), y = 25000, label = "Cumulative deaths")

ggplotly(p)
```

We will be looking at the COVID-19 

# How life was affected in Iraq? What does the data tell us?

You have read a long 

```{r}
iraq_demographics <- read.csv(here::here("Blog_entry_1/data", "iraq_demographics.csv"))
```


# Tribute to factfullness

The world is a better place and you just have to believe it
 





# Conclusion


# Fun fact
For this exercise, i again encountered 

# Reference
Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686, https://doi.org/10.21105/joss.01686.  
Winston Chang, Joe Cheng, JJ Allaire, Yihui Xie and Jonathan McPherson (2020). shiny: Web Application Framework for R. R package version 1.4.0.2.https://CRAN.R-project.org/package=shiny.  
Joe Cheng, Bhaskar Karambelkar and Yihui Xie (2019). leaflet: Create Interactive Web Maps with the JavaScript 'Leaflet' Library. R package version 2.0.3. https://CRAN.R-project.org/package=leaflet  
Yihui Xie, Joe Cheng and Xianying Tan (2020). DT: A Wrapper of the JavaScript Library 'DataTables'. R package version 0.15. https://CRAN.R-project.org/package=DT  
H. Wickham. ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag New York, 2016.  
Thomas Lin Pedersen and David Robinson (2020). gganimate: A Grammar of Animated Graphics. R package version 1.0.5. https://CRAN.R-project.org/package=gganimate  
C. Sievert. Interactive Web-Based Data Visualization with R, plotly, and shiny. Chapman and Hall/CRC Florida, 2020.  
Erich Neuwirth (2014). RColorBrewer: ColorBrewer Palettes. R package version 1.1-2. https://CRAN.R-project.org/package=RColorBrewer  
Hao Zhu (2019). kableExtra: Construct Complex Table with 'kable' and Pipe Syntax. R package version 1.1.0. https://CRAN.R-project.org/package=kableExtra  
Garrett Grolemund, Hadley Wickham (2011). Dates and Times Made Easy with lubridate. Journal of Statistical Software, 40(3), 1-25. URL http://www.jstatsoft.org/v40/i03/.  
Jeroen Ooms (2018). gifski: Highest Quality GIF Encoder. R package version 0.8.6. https://CRAN.R-project.org/package=gifski  
Simon Urbanek (2013). png: Read and write PNG images. R package version 0.1-7. https://CRAN.R-project.org/package=png  
Jeroen Ooms (2020). magick: Advanced Graphics and Image-Processing in R. R package version 2.4.0. https://CRAN.R-project.org/package=magick  
Alburez-Gutierrez, D. (2020). Population pyramids in R for beginners. Retrieved 3 September 2020, from http://alburez.me/2018-03-20-Population-pyramids-in-R-for-beginners/  
Download today’s data on the geographic distribution of COVID-19 cases worldwide. (2020). Retrieved 2 September 2020, from https://www.ecdc.europa.eu/en/publications-data/download-todays-data-geographic-distribution-covid-19-cases-worldwide  
Haverty, Mironova, Hlaing, Fishbein, Garrett, & Mitnick et al. (2020). Israel’s Cautionary Coronavirus Tale. Retrieved 4 September 2020, from https://foreignpolicy.com/2020/07/22/israel-coronavirus-second-wave-netanyahu/  
Israel Geographic coordinates - Latitude & longitude. (2020). Retrieved 4 September 2020, from https://www.geodatos.net/en/coordinates/israel  
Israel latitude and longitude. (2020). Retrieved 4 September 2020, from https://latitudelongitude.org/il/  
Sandri, M. (2020). gganimate time series and two line plot. Retrieved 3 September 2020, from https://stackoverflow.com/questions/54855334/gganimate-time-series-and-two-line-plot  
https://www.ifimes.org/en/7991-the-current-situation-in-iraq

https://www.bbc.com/news/world-middle-east-14546763

https://www.wmf.org/project/cultural-heritage-sites-iraq

https://en.wikipedia.org/wiki/PUK_insurgency

https://uca.edu/politicalscience/dadm-project/middle-eastnorth-africapersian-gulf-region/iraqkurds-1932-present/

https://link.springer.com/content/pdf/bbm%3A978-1-137-60181-0%2F1.pdf

https://stackoverflow.com/questions/43093712/draw-a-map-of-a-specific-country-with-leaflet
 

“ACLED. (2030). “Armed Conflict Location & Event Data Project (ACLED) Codebook, 2020.””

“Raleigh, Clionadh, Andrew Linke, Håvard Hegre and Joakim Karlsen. (2010). “Introducing ACLED-Armed Conflict Location and Event Data.” Journal of Peace Research 47(5) 651-660.”

https://github.com/timelyportfolio/leaftime/issues

