---
title: "How are the Jewish doing?"
description: |
  Let's dwelve into our first blog by doing a small visualisation exercise of the covid situation in Israel! I intend to make this first blog to be less formal but more friendly, so we can get to know each other better. Let's visualize together!
author:
  - name: Hanh Ngo
    ID: 31196101
    institution: Monash University
    program: Master of Business Analytics
date: 09-12-2020
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
library(tidyverse)
library(shiny)
library(leaflet)
library(DT)
library(RColorBrewer)
library(gganimate)
library(ggplot2)
library(kableExtra)
library(lubridate)
library(plotly)
library(gifski)
library(png)
library(magick)
```

Hello! and welcome to my first blog!

Let's get to know each other by doing a little bit of exercise. Today i will be looking at the most heated topics being mentioned recently - the Corona virus. To be specific, i look at the Covid situation in one country of our choice, using all of the tools i have learnt so far trying to see if there is anything i can learnt from the dataset. The country we will be looking at today is Israel. 

# Data description

For this exercise, i used a dataset provided from European Centre for Disease Prevention and Control (ECDC) which can be obtained [here](https://www.ecdc.europa.eu/en/publications-data/download-todays-data-geographic-distribution-covid-19-cases-worldwide) for the information of the number of cases and deaths per day recording from 31 Dec 2019 up to the most recent (3 September 2020 by the day of this blog post).

Other information including the locations of cases, the age dispersion and the hospital status was obtained directly from the Israel Ministry of Health (MOH)'s [daily dashboard](https://datadashboard.health.gov.il/COVID-19/general) and [Telegram channel](m https://t.me/MOHrepor). 

The process of obtaining data was not easy as even though the Israel Government kept the public up to date with timely update and release of data and figures, most of them were published in Hebrew. The trace for data started from the [Wikipedia page](https://en.wikipedia.org/wiki/Template:COVID-19_pandemic_data/Israel_medical_cases#cite_note-1) where i obtained the offical data source of the MOH, then followed along a tidious work of tracking down the MOH's telegram and downloading every files offered since i did not know which was which. After the extensive use of google translate and googling (to verify the translation work) to decide what files to use comes the even more tedious works of manually obtaining the GPS coordinates for each Israel cities appreared in the dataset for the mapping job later. It was bit a tricky that each cities written in Hebrew could yield quite some names in English due to the pronounciation difference, thus it was no easy job to match the name of the cities in the dataset with the ones in the GPS services. I would be happy if anyone could suggest their solutions for a shortcut or a roundway for all of the manual works that i had carried for this mini project. 

After some days of data digging, understanding and cleaning, we are ready for the analysis.

# The overall situation

## What is the situation as of now?

```{r data-load}
data <- read.csv("https://opendata.ecdc.europa.eu/covid19/casedistribution/csv", na.strings = "", fileEncoding = "UTF-8-BOM")
raw_israel <- data %>%
  filter(countriesAndTerritories == "Israel") 
```

```{r raw-data}
raw_dat <- raw_israel %>%
  select(dateRep, day, month, year, cases, deaths, popData2019) %>%
  arrange(year, month, day) %>%
  mutate(cumulative_case = cumsum(cases),
         cumulative_death = cumsum(deaths)) 
```

Let's have a look at the current situation in Israel in the most recent week up to the 3 September 2020:

```{r summary-stats}
summary <- raw_dat %>%
  # add separator
  mutate_if(is.numeric, funs(prettyNum(., big.mark=","))) %>%
  filter (dateRep %in% c("28/08/2020","29/08/2020", "30/08/2020", "31/08/2020", "01/09/2020",
                         "02/09/2020","03/09/2020")) %>%
  select(- year,- month, - day, - popData2019)
# Gather additional info off the Gov's telegram
add_info <- tibble::tribble(
  ~active_patient, ~recovered, ~test_performed_inday,
  20055, 91051, 36371,
  20331, 92100, 23263,
  20305, 92796, 10592,
  20648, 95009, 13406,
  20960, 96206, 31705,
  23261, 97234, 28259,
  24281, 98637, 34857
) %>%
  # add separator
  mutate_each(funs(prettyNum(., big.mark=",")))

# cbind all the columns
summary <- summary %>%
  cbind(add_info)
# print off the table
summary %>%
  kable (caption = "Summary of 7 days") %>%
  kable_styling(bootstrap_options = c("hover", "stripped")) %>%
  row_spec(7, bold = T, background = "yellow")
```


The highlighted row indicated the the most recent day presented in the dataset. The country has a total of 122,539 confirmed cases with 969 deaths. The death rate accounted roughly 0.8% of total cases, highlighting the Government's achivement to handle the situation. 

Please note that eventhough this post was made in the 3 Sep 2020, the dataset would be updated daily which would shift the 7 days presented here onwards. For the purpose of this report only, i chosed to present specificly the 7 days prior to 3 Sep 2020 and did not allow the data to be updated in the table. The reason for this was that the original dataset obtained from the ecdc did not include the information for active patient, number of recover and number of test performed daily. These information was obtained from the Govermental daily report and added manually to the table. As such, to avoid conflicts in the data presented, i fixed the 7 days presented for the sake of this report only. 

## The timeline
Next up we will have a look at the change by days of the pandemic in Israel to track down the start of the breakout. 

```{r change-by-day}
# Stack data to prepare for plotting
dat2 <- raw_dat %>%
  select(dateRep, cumulative_case, cumulative_death) %>%
  mutate(dateRep = dmy(dateRep))

dat2 <- data.frame(Date=rep(dat2$dateRep, 2), 
                  case_death=c(dat2$cumulative_case, dat2$cumulative_death), 
                  grp=rep(c("case","death"), each=nrow(dat2)))
# animation
p <- ggplot(dat2, aes(x=Date, y=case_death, group=grp)) +
  geom_line(color = "blue") +
  geom_segment(aes(xend=max(Date), yend = case_death), linetype=2, colour='#CC0033') +
  geom_point(size = 3, colour = "#0000FF") + 
  geom_text(aes(x = max(Date)+.1, label = sprintf("%5.0f", case_death)), hjust=0) +
  transition_reveal(Date) + 
  coord_cartesian(clip = 'off') + 
  view_follow(fixed_y = TRUE) +
  labs(title = 'Change by days', y = 'Number of cases (upper) / death (lower)') +
  enter_drift(x_mod = -1) + exit_drift(x_mod = 1) +
  theme_bw() +
  theme(panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"),
        plot.margin = margin(5.5, 40, 5.5, 5.5))

   animate(p, fps=5)
```

Israel did not record the first case until the end of February when a woman was confirmed positive after returning from the _Diamond Princess_ cruise ship in Japan. The first deased was confirmed in March 2020 of an 88-year-old man in Jerusalem.

The Government was quick to response to the pandemic when on 9 March, the Government had requested 14-days isolation for all of whom who travelled to Israel. Quickly on 18 March, the country started to close off the borders with the world, just a couple of weeks before the outbreak in April. Early responds coupled with efficient health care system, strict curfews and policy had nominated Israel as one of the early success face in the battle against the novel virus. As you can see from the chart, the curve was almost flattened in the first months with a mild incresed around April to June. 

In July, the country faced the second wave of which created a significant impact when a whooping 44 thousand cases were confirmed in just one month, compared to 24 thousand cases recored from March to June combined. Ironically, the reason for such outbreak stemmed from the success of the first wave which gave a false "sense of assurance" and knocked down the country's guard. Early school reopening in May coupled with the abandon of social distancing policy and hybrid learning between online and on-campus had spiked up the number of cases in the upcoming months, raising from 24 thousand cases in the first 6 months to more than 122 thousand in the next 3 months. This lesson was learnt in such a hard and bitter way. 

# Location 

Next up i attempted to explore the dispersion of cases accross the country. I attempted to map as much locations as possible. There were a few that i chosed to exclude, which were locations that had lower than 15 confirmed cases and those which i could not obtained the coordinated. The map below was the result of such attempt.

This map, however did not show the number of confirmed cases by area, instead it was the number of active patients that was shown. The reason was simply  because i could only obtain the "active patients" by area report provided by the Government.

```{r city-dat}
# read the location
city_raw <- read.csv(here::here("Blog_entry_1", "israel_city_02_09_20.csv"))
city <- city_raw %>%
  select(city_ENG, active_case, lat, lon) %>%
  # Pick only location with more than 15 active cases
  filter(!active_case == "Less than 15",
  # Filter out locations without coordinates
         !is.na(lat)) %>%
  mutate(active_case = as.numeric(as.character(active_case)),
    # categorize for plotting purpose
    category = case_when(
    active_case <= 50 ~ "<= 50 cases",
    active_case > 50 & active_case <= 100 ~ "51 - 100 cases",
    active_case > 100 & active_case <= 500 ~ "101 - 500 cases",
    active_case > 500 & active_case <= 1000 ~ "501 - 1000 cases",
    TRUE ~ "more than 1000 cases")
  )
```

```{r map-location}
# Create factor for legend
case_factor <- factor(city$category, 
                      levels = c("<= 50 cases","51 - 100 cases","101 - 500 cases","501 - 1000 cases","more than 1000 cases"),
                      labels = c("<= 50 cases","51 - 100 cases","101 - 500 cases","501 - 1000 cases","more than 1000 cases"))

# Create colour palette
pal <- colorFactor(pal = "YlOrRd", 
                   levels = case_factor, reverse = TRUE)

# Draw leaflet
leaflet() %>%
addProviderTiles(providers$Stamen.Toner) %>%
setView(34.4, 33.2, 7) %>%
addCircleMarkers(data = city, lat =  ~lat, lng =~ lon,
           popup = ~as.character(paste0('<strong>City/ Town: </strong>',city_ENG,
                                          '<br><strong>Active case:</strong> ', active_case)),
           color = ~pal(city$category),
           radius = 3, fillOpacity = 0.8, stroke = FALSE) %>%
  addLegend(pal = pal, values = case_factor, opacity=1,
            title = "No of cases")
```

The dots presented in the map represented for each city in the dataset and was coloured based on the number of cases. The yellow and light orange dots dominated the map indicating that each city/ regional area is doing a wonderful job keeping the situation under control. The high-case area cluttered around the Tel Aviv area with the highest records belonged to Jerusalem (2601 active patients). These area are the metropolitan cities with high people density. Most cases cluttered near the shoreline while the in-land area was mostly "white" out. 

# Age dispersion

A quick tour at the age dispersion of the patients in this country showed sad facts that could not be emphasized enough. 

```{r death-age}
age_raw <- read.csv(here::here("Blog_entry_1", "death_report_02_09_20.csv"))
age_death <- age_raw %>%
  select(4:5) %>%
  filter(!gender == "unknown") %>%
  mutate(age_group = case_when(
    age >= 0 & age <10 ~ "0 - 9",
    age > 9 & age < 20 ~ "10 - 19",
    age > 19 & age < 29 ~ "20 - 29",
    age > 29 & age < 39 ~ "30 - 39",
    age > 39 & age < 49 ~ "40 - 49",
    age > 49 & age < 59 ~ "50 - 59",
    age > 59 & age < 69 ~ "60 - 69",
    age > 69 & age < 79 ~ "70 - 79",
    age > 79 & age < 89 ~ "80 - 89",
    TRUE ~ "90+")
  ) %>%
  group_by(age_group, gender) %>%
  count() %>%
  # flip female ages to negative side of the chart
  mutate(prop = n/(nrow(age_raw)-1)*100,
         prop = ifelse(gender=="female", prop*(-1), prop*1)) %>%
  ungroup()

p2 <- age_death %>%
ggplot(aes(x = age_group, y = prop, fill = gender)) + 
  geom_bar(data = subset(age_death, gender == "female"), stat = "identity") + 
  geom_bar(data = subset(age_death, gender == "male"), stat = "identity") + 
  scale_y_continuous("percentage (%)", breaks = scales::breaks_pretty(n = 6),
                     # Convert negative number to positive
                     labels = function(br) ifelse(br>=0,br, abs(br))) + 
  coord_flip() + 
  scale_fill_brewer(palette = "Set1") + 
  xlab("") +
  theme_classic() +
  theme(legend.position = "none")
```

```{r case-age, layout="l-body-outset", preview = TRUE}
# Percentage from gov databoard for 2 Sep 2020
age_case <- tibble::tribble(
  ~age_group, ~gender, ~percentage,
  "0-9", "female", 5.20,
  "10-19", "female", 8.50,
  "20-29", "female", 10.30,
  "30-39", "female", 7,
  "40-49", "female", 6.60,
  "50-59", "female", 5.10,
  "60-69", "female", 3.20,
  "70-79", "female", 1.60,
  "80-89", "female", 1,
  "90+", "female", 0.5,
  "0-9", "male", 5.30,
  "10-19", "male", 11,
  "20-29", "male", 10.5,
  "30-39", "male", 7.10,
  "40-49", "male", 6.20,
  "50-59", "male", 4.90,
  "60-69", "male", 3.40,
  "70-79", "male", 1.70,
  "80-89", "male", 0.70,
  "90+", "male", 0.20) %>%
  # flip female ages to negative side of the chart 
  mutate(percentage = ifelse(gender=="female", percentage*(-1), percentage*1))

p3 <- age_case %>%
ggplot(aes(x = age_group, y = percentage, fill = gender)) + 
  geom_bar(data = subset(age_case, gender == "female"), stat = "identity") + 
  geom_bar(data = subset(age_case, gender == "male"), stat = "identity") + 
  scale_y_continuous("percentage (%)", breaks = scales::breaks_pretty(n = 6),
                     # Convert negative number to positive
                     labels = function(br) ifelse(br>=0,br, abs(br))) + 
  coord_flip() + 
  scale_fill_brewer(palette = "Set1") + 
  xlab("") +
  labs(title = "Death (left) and Case (right) proportion by age group and gender") +
  theme_classic()

# Put two plots side by side
subplot(p2,p3)
```

The right population pyramid illustrated the age dispersion of the confirmed_cases by gender and age group in proportion instead of absolute number. Clearly the youngs were the most exposed to the virus with this pyramid having a fat bottom. The proportion of infected gender roughly equalled with male having slightly more cases than female. However, moving to the second chart of the death by genders and age group, it's sad to see that eventhough accounted for a small number of confirmed cases, the seniors were the most vulnerable. Most deaths were recorded at the age of 80 and above with male in their 70s had higher death cases than female. However, female at the later age (the 90s) were weaker to the disease as 18% of the deaths coming from this group, compared to 15% of male.

Hopefully this finding could again serve as a warning for all of us young and fun out there to know that all the strict policies and protective action are meant to protect our grandparents and our parents, the ones who are most vulnerable to this virus.

# Hospital situation

This exercise will conclude with a small report of the hospital status as of 3 March 2020 in Israel. The list below can be extracted to various format, containing the status of the 30 biggest hospitals in Israel monitored closely by the Government. The name in English was roughly translated using google so i would be appreciated if anyone knowing Hebrew can helo me fix if i made any errors in translation. 

```{r}
hospital_status <- read.csv(here::here("Blog_entry_1", "hospital_status.csv")) %>%
  select(1:5)

# Print the hospital table
datatable(hospital_status, filter = "top",  
            extensions =  'Buttons', 
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                           lengthMenu = list(c(10,20,30,-1),c(10,20,30,"All"))
                           ),
            colnames = c("Hospital_Hebrew", "Hospital_Eng", "Area", "Occupancy", "Corona_Occupancy")
            )
```

We can see that mose hospitals are starting to reach the maximum capacity with some has already exceed their limits. The proportion of capacity spent for battling the virus was also recorded at the Corona_Occupancy column. If you are happen to be in Israel at the moment, i hope this list can help you locate the hospitals to avoid to reduce the workloads for the doctors and nurses working there. If you are not in Israel? then i hope it's a nice FYI to finish this mini exercise. 

# Conclusion
This exercise was a fun bit to do. We learnt how our friends in Israel are doing, how the situation there, and most importanly after this exercise hopefully we can learn to read a bit of Hebrew :D The overall practice was not tremendously difficult as we only made use of the familiar ggplot with supports from plotply, leaflet and gganimate to make our post more colourful. The most challenging and exciting parts came from the data mining and digging process, which mostly due to the language barrier. I hope you are enjoy reading this blog and any comments to improve the contents are warmly welcomed.


# Reference
Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686, https://doi.org/10.21105/joss.01686.  
Winston Chang, Joe Cheng, JJ Allaire, Yihui Xie and Jonathan McPherson (2020). shiny: Web Application Framework for R. R package version 1.4.0.2.https://CRAN.R-project.org/package=shiny.  
Joe Cheng, Bhaskar Karambelkar and Yihui Xie (2019). leaflet: Create Interactive Web Maps with the JavaScript 'Leaflet' Library. R package version 2.0.3. https://CRAN.R-project.org/package=leaflet  
Yihui Xie, Joe Cheng and Xianying Tan (2020). DT: A Wrapper of the JavaScript Library 'DataTables'. R package version 0.15. https://CRAN.R-project.org/package=DT  
H. Wickham. ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag New York, 2016.  
Thomas Lin Pedersen and David Robinson (2020). gganimate: A Grammar of Animated Graphics. R package version 1.0.5. https://CRAN.R-project.org/package=gganimate  
C. Sievert. Interactive Web-Based Data Visualization with R, plotly, and shiny. Chapman and Hall/CRC Florida, 2020.  
Erich Neuwirth (2014). RColorBrewer: ColorBrewer Palettes. R package version 1.1-2. https://CRAN.R-project.org/package=RColorBrewer  
Hao Zhu (2019). kableExtra: Construct Complex Table with 'kable' and Pipe Syntax. R package version 1.1.0. https://CRAN.R-project.org/package=kableExtra  
Garrett Grolemund, Hadley Wickham (2011). Dates and Times Made Easy with lubridate. Journal of Statistical Software, 40(3), 1-25. URL http://www.jstatsoft.org/v40/i03/.  
Jeroen Ooms (2018). gifski: Highest Quality GIF Encoder. R package version 0.8.6. https://CRAN.R-project.org/package=gifski  
Simon Urbanek (2013). png: Read and write PNG images. R package version 0.1-7. https://CRAN.R-project.org/package=png  
Jeroen Ooms (2020). magick: Advanced Graphics and Image-Processing in R. R package version 2.4.0. https://CRAN.R-project.org/package=magick  
Alburez-Gutierrez, D. (2020). Population pyramids in R for beginners. Retrieved 3 September 2020, from http://alburez.me/2018-03-20-Population-pyramids-in-R-for-beginners/  
Download today’s data on the geographic distribution of COVID-19 cases worldwide. (2020). Retrieved 2 September 2020, from https://www.ecdc.europa.eu/en/publications-data/download-todays-data-geographic-distribution-covid-19-cases-worldwide  
Haverty, Mironova, Hlaing, Fishbein, Garrett, & Mitnick et al. (2020). Israel’s Cautionary Coronavirus Tale. Retrieved 4 September 2020, from https://foreignpolicy.com/2020/07/22/israel-coronavirus-second-wave-netanyahu/  
Israel Geographic coordinates - Latitude & longitude. (2020). Retrieved 4 September 2020, from https://www.geodatos.net/en/coordinates/israel  
Israel latitude and longitude. (2020). Retrieved 4 September 2020, from https://latitudelongitude.org/il/  
Sandri, M. (2020). gganimate time series and two line plot. Retrieved 3 September 2020, from https://stackoverflow.com/questions/54855334/gganimate-time-series-and-two-line-plot  
קורונה - לוח בקרה. (2020). Retrieved 4 September 2020, from https://datadashboard.health.gov.il/COVID-19/general  
קורונה - משרד הבריאות. (2020). Retrieved 4 September 2020, from https://t.me/MOHreport  
COVID-19 pandemic data/Israel medical cases. (2020). Retrieved 4 September 2020, from https://en.wikipedia.org/wiki/Template:COVID-19_pandemic_data/Israel_medical_cases#cite_note-1  
The Novel Coronavirus - Israel Ministry of Health. (2020). Retrieved 4 September 2020, from https://govextra.gov.il/ministry-of-health/corona/corona-virus-en/  


 



