---
title: "Unboxing an R package"
description: |
  Unboxing stuffs has always been a pleasure, and it would be even more so when you spent hours choosing the items inside or actually building the items inside. I am very excited today to unbox an R package created by one of my classmate for his assignment. We will have a look at his package, trying to learn what he had done best and what he could do to improve his work.  
author:
  - name: Hanh Ngo
    affiliation: Monash University
    program: Master of Business Analytics
date: 10-28-2020
output:
  distill::distill_article:
    self_contained: false
---

The package I will be reviewing today was created by Dea Avega Editya from my Master class and named **covidasean**. The **covidasean** package is a container for a Shiny dashboard Dea made for his previous assignment reporting the COVID situation in the ASEAN region.  

# An app overview

The covidasean package could be downloaded and installed via Dea's github [here](https://github.com/etc5523-2020/r-package-assessment-dedi0003). I was able to install and build the package locally without any problems.

To launch the app contained in the package, Dea set up a function called `launch_app()` which can be typed into the console to call the app out. A minor problem arised when two of the dependence packages he used for his Shiny app, the _tidycovid19_ and _wbstats_, were temporarily taken off CRAN. Users who did not have these two packages installed beforehand have to install them remotely from Github. Eventhough it took extra works, the app launched successfully after the two missing packages were installed.

The package's structure contained everythings we need to build and run a simple package. As we can see from an image below, Dea had created a description, a readme, documentation for the main two of his five R functions (nested under the _man_ folder), vignettes and two unit tests for his package which is enough to get you started with. Moreover, his [pkgdown page](https://etc5523-2020.github.io/r-package-assessment-dedi0003/) was promptly set up which really helps navigating through his package. Well done here Dea!

```{r eval = FALSE}
fs::dir_tree(path = "/Users/HanseMac/Desktop/Monash/ETC5523_Communicating_with_data/Blog_entry/r-package-assessment-dedi0003-master", recurse = FALSE)
```

```{r structure}
# paste pictures to avoid path conflicts
knitr::include_graphics(file.path(getwd(),"structure.png"))
```

# Potential area for improvement

Whilst Dea was doing a wonderful job creating his first R package, there certainly are a lot of areas for improvement. I will divide my recommendation into two separate parts: the package and the Shiny app. 

## The package

**1. The package structure**

If you noticed, Dea had a **covidsean** folder existed inside his R project. I was wondering what that was at first. Turns out it is a sub folder that contains the exact everything that the parent folder has. My best guess was that eventhough the original class repo was created as a package already, when Dea cloned the repo down to start building his package, he used `usethis::create_package("pkgname")` to yet again create another package nested inside a package. I figured Dea could have faced a lot of troubles for this. I would totally recommend him removing all the unused folders/ files to avoid clustering and confusion for others.  

**2. The github pages**

Whilst I am happy that his pkgdown page has done me a great favour guiding me through the app, I must confess that it took me a while to locate where it resides. Normally if we specified the source for the Github pages in the Setting as below:

```{r github-setting}
knitr::include_graphics(file.path(getwd(),"page-setting.png"))
```

On the right hand side of the "Code" page, next to the Readme, you would find something under the Environments. Clicking on the **github-pages** would bring you to the action log where you can get access to the pkgdown site by viewing deployment. 

```{r github-page, out.height= '50%'}
knitr::include_graphics(file.path(getwd(),"page.png"))
```

I manages to track down Dea's pkgdown site by visiting the chat he had with our lecturer in the Pull requests tab. I would recommend Dea to set up the Github pages source for easier access to his pkgdown site.

**3. The unit tests**

Dea had two unit tests for his package. One is for testing the length of the `render_case()` function which was expected to return 1 for only one plot generated; the other test is for the length of the `left_sidebar function which would expect to return 3 for three inputs he set up in the panel. All tests passed smoothly.

I would say whilst I'm happy that the tests worked for him (i'm sure everyone knew it could be stressful get a pass on a unit test when your deadline drawed near), i think these two tests are a little bit simple and somehow obvious to me. Further more, there is not much to expand for these two tests as the object tests are quite simple. I would suggest there are other potential areas to build the unit test on, for example, his table. 

Dea set up a table that would return results based on your selected date input. However, there is a latent error embedded in this kind of input, that what would happen if your selected dates are out of the range of the dataset employed? It would return an error like in the picture below:

```{r}
knitr::include_graphics(file.path(getwd(),"date-error.png"))
```

There are two noticable criteria for this date input to work:

- Both the begin date and end date have to be smaller than the `max(date)` of the dataset.  
- The end date has to be bigger than the begin date.  

I at first did not notice but when i started to scan out for potential candidate for the unit test, this glitch was brought to attention. It is easy to neglect points like this, especially when you work with a lot of inputs and functions, spanding across a lot of pages in your dashboard.

## The Shiny app

**1. The code**

The highlight of the dashboard was the plotly-click charts where, in this case, you can click on each country presented in the confirmed cases chart (the blue bar chart) and subsequently, the line plot underneath will show the corresponding trend lines (the line chart) for your selected country as shown in the extract below.

```{r}
knitr::include_graphics(file.path(getwd(),"click-error.gif"))
```

However, as you can see, there was an error produced when you click on the line instead of the bar as instructed in the guidance. In fact, such error would also occur if you click on the lines in the "Trend of People's Mobility" chart. So what happended here?

First we will have a look at the source code in the belowing code block:

```{r example, eval = FALSE, echo = TRUE}
# Code to produce bar chart
    output$case <- covidasean:::render_case() 
    
# Code to produce line chart
    output$trendchart <- renderPlotly({
        d <- event_data("plotly_click")
            if (is.null(d) return(
                case_plot <- covid_asean_df %>%
                    filter(country == "Indonesia") %>%
                    ggplot(aes(x = date)) +
                    geom_line(aes(y = ave_new_cases, color ="avg cases"), size = 1) +
                    geom_line(aes(y = ave_new_deaths, color ="avg deaths"), size = 1, alpha = 0.5) +
                    geom_line(aes(y = ave_new_recovered, color ="avg recovered"), size = 1, alpha =0.5) +
                    theme_minimal() +
                    ylab("Count")+
                    ggtitle(paste0("Covid-19 Cases in Indonesia", d$x))+
                    theme(panel.grid.major.x = element_line(),
                          panel.grid.minor = element_blank(),
                          legend.position="top", legend.box = "horizontal")
            )

        case_plot <- covid_asean_df %>%
            filter(country %in% d$x) %>%
            ggplot(aes(x = date)) +
            geom_line(aes(y = ave_new_cases, color ="avg cases"), size = 1) +
            geom_line(aes(y = ave_new_deaths, color ="avg deaths"), size = 1, alpha = 0.5) +
            geom_line(aes(y = ave_new_recovered, color ="avg recovered"), size = 1, alpha =0.5) +
            theme_minimal() +
            ylab("Count")+
            ggtitle(paste0("Covid-19 Cases in ", d$x))+
            theme(panel.grid.major.x = element_line(),
                  panel.grid.minor = element_blank(),
                  legend.position="top", legend.box = "horizontal")

            return(ggplotly(case_plot)) %>%
                config(displayModeBar = F)  %>%
                layout(legend = list(orientation = "h", x = 0.2, y = -0.1))
    })
```

He had the bar chart created by running a function from an external R script called `render_case.R` as, i believed, an attempt to refactor the code. We should check that out too:

```{r eval = FALSE, echo = TRUE}
render_case <- function(){
renderPlotly({
  total_case <- covid_asean_df %>%
    group_by(country) %>%
    summarise(confirmed_cases = sum(new_cases))

  total_case %>%
    plot_ly(x = ~country, y = ~confirmed_cases) %>%
    config(displayModeBar = F)
}) }
```

There are three problems from these two code blocks:

- Dea had not specified the _"source"_ plot which the _"plotly_click"_ event referred to. As a result, whenever Shiny detects a _"plotly_click"_ event, it will change the plot, disregarding of the fact that wrong "clicked" plot will lead to the event plot not displaying correctly. In this case, the source plot is the bar plot whilst the event plot is the "COVID-19 Cases in xxx" line plot.   
- Fixing this error is not complicated, however the way Dea structured the code made the fix struggling. The source plot in the app was created by calling out a function. However, the _"plotly_click"_ event might not recognise the _"source"_ identified for a plot put within a function from an external source (i.e not in the same R script with an event plot).  
- The code to produce the line chart is quite lengthy and is a potential candidate for refactoring. Dea's idea was that when there is no _"plotly_click"_ event, the default chart will be of Indonesia, not any other countries.    

Whilst his ideas are great, we can definitely make them better. I had tried to improvise his code, addressing the mentioned problems. The result is shown below:

```{r}
knitr::include_graphics(file.path(getwd(),"fix-error.gif"))
```

Here is the improvised source code, I have also included comments to places where i made ammendment. 

```{r improvision, eval = FALSE, echo = TRUE}
# Code to build bar chart
output$case <- 
    # Need to bring the plot to the same place with plotly_click event to trigger the source
        renderPlotly({
            total_case <- covid_asean_df %>%
                group_by(country) %>%
                summarise(confirmed_cases = sum(new_cases))
            p <- total_case %>%
              # Indicate the source
                plot_ly(x = ~country, y = ~confirmed_cases, source = "to-click") %>%
                config(displayModeBar = F)
            p })

    # Create this new variable to response to the plotly_click event
    selectedcountry <- reactiveVal()
    observeEvent(event_data("plotly_click", source = "to-click"), {
        selectedcountry(as.character(event_data("plotly_click", source = "to-click")$x))
    })
    
# Code to build line chart
output$trendchart <- renderPlotly({
    # Create display_country to help shorten the code and remove the Indonesia code block
      if (is.null(selectedcountry())) {
      display_country <-  "Indonesia"}
      else {display_country <- selectedcountry()}
  
    # Remove entire Indonesia code block
           # return(
            #    case_plot <- covid_asean_df %>%
            #        filter(country == "Indonesia") %>%
            #        ggplot(aes(x = date)) +
            #        geom_line(aes(y = ave_new_cases, color ="avg cases"), size = 1) +
            #        geom_line(aes(y = ave_new_deaths, color ="avg deaths"), size = 1, alpha = 0.5) +
                    # geom_line(aes(y = ave_work, color ="Workplaces"), size = 1.2) +
                    # geom_line(aes(y = ave_station, color ="Stations"), size = 1.2)+
            #        geom_line(aes(y = ave_new_recovered, color ="avg recovered"), size = 1, alpha =0.5) +
            #        theme_minimal() +
            #        ylab("Count")+
            #        ggtitle(paste0("Covid-19 Cases in Indonesia", selectedcountry()))+
                    # facet_wrap(.~ country, nrow = 1)+
            #        theme(panel.grid.major.x = element_line(),
            #              panel.grid.minor = element_blank(),
            #              legend.position="top", legend.box = "horizontal")
            #)

      case_plot <- covid_asean_df %>%
        # Replace the display_country in the filter
            filter(country %in% display_country) %>%
            ggplot(aes(x = date)) +
            geom_line(aes(y = ave_new_cases, color ="avg cases"), size = 1) +
            geom_line(aes(y = ave_new_deaths, color ="avg deaths"), size = 1, alpha = 0.5) +
            geom_line(aes(y = ave_new_recovered, color ="avg recovered"), size = 1, alpha =0.5) +
            theme_minimal() +
            ylab("Count")+
        # Replace the display_country also here to create the title
            ggtitle(paste0("Covid-19 Cases in ", display_country))+
            theme(panel.grid.major.x = element_line(),
                  panel.grid.minor = element_blank(),
                  legend.position="top", legend.box = "horizontal")
            return(ggplotly(case_plot)) %>%
                config(displayModeBar = F)  %>%
                layout(legend = list(orientation = "h", x = 0.2, y = -0.1))
    })
```


**2. The app layout**

The layout of the plot is a bit confusing to me. In the side bar we have three inputs to choose from the list. However, the first two belongs to the Mobility plot which located at the bottom of the page - furthest from the input panel. The last one had nothing to do with either of the plots but the tables located inside the side bar. 

I was also first at lost of the Bibliography button when i could not locate the Bibliography when i chose "Show". Turned out the unfolded content was put way in the bottom of the page, which you have to scroll down to see. It was hard to catch if you did not look carefully. 

You can have a glimpse through the app interface in the screenshot below:

```{r dashboard}
knitr::include_graphics(file.path(getwd(),"dashboard.png"))
```

To make the layout friendlier, I would recommend Dea to:    

- Separate the inputs into different, distinguised parts of the side bar rather than put all of them together in one place. Put the plots/ tables related to the inputs closed to each other.  
- The "select country" inputs is somehow related to the plotly_event click plots where "country" served as a trigger for change. I would suggest link the table and these first two plots together by using `observeEvent` combined with `updateSelectizeInput`. That way, when ones chose a country from a drop down list, the table and the "Covid-19 Cases" line plot will change at once. Similary, when one chosed a country from the bar plot, the table and the "Covid-19 Cases" line plot will also change. In my Shiny app, i was able to perform this combo but it will take quite a handful of trials and errors to finally get it work. I will try to get my hands on Dea's app to reflect my recommendation when i have more time.   
- I would recommend to put the Bibliography (or References) and the introduction of the app into a separate page to avoid clustering.  

# Tips to bag
• A paragraph wrapping up what you’ve learned from reviewing the package? (2 marks)

# Appendix

3. Complete the check list on the following page that reviews different components of the package and add it as an appendix to your post. (5 marks) 

## Package Review
*Please check off boxes as applicable, and elaborate in comments below.*

#### Documentation  
The package includes all the following forms of documentation:  

- [x] **Installation instructions:** for the package is found in README  
- [ ] **Vignette(s)** demonstrating major functionality that runs successfully locally  
- [x] **Function Documentation:** for all exported functions in R help  
- [ ] **Examples** for all exported functions in R Help that run successfully locally 

#### Functionality  
- [x] **Installation:** Installation succeeds as documented.  
- [ ] **Functionality:** Any functional claims of the software been confirmed.  
- [ ] **Automated tests:** Unit tests cover essential functions of the package and a reasonable range of inputs and conditions. All tests pass on the local machine.

